---
title: "Practical Machine Learning Project AVR"
author: "AVR"
date: "November 20, 2018"
output:
  html_document: default
  pdf_document: default
---
##EXECUTIVE SUMMARY
The goal of this project is to use the data of 6 participants from accelerometers on the belt, forearm, arm and dumbell predict the manner in which they did the exeercise (they were asked to perform barbell lifts correctly and incorrectly in 5 different ways). What to submit : a report should be created to describe how the model was built, how cross validation was used, what is expected out of the sample error and why were the choices made. In addition, Prediction model will be used to predict 20 different cases. Only less than 2000 words and 5 or less figures will be used.

##DATA
The data to be used are : (training) https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv and (test) https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load R Packages, read and clean data.  Split the data set to 70-30.
```{r Packages Installed}
library(caret)
library(randomForest)
library(corrplot)
set.seed(2018)

setwd("D:/DOST Coursera/Module8")
training_data = read.csv("pml-training.csv", na.strings=c("", "NA"))

training_data = training_data[8:length(training_data)]
cln = colSums(is.na(training_data))
training_data = training_data[,cln == 0]

inTrain = createDataPartition(training_data$classe, p = .7)[[1]]
training = training_data[inTrain,]
validation = training_data[-inTrain,]
```
## Plot correlation matrix
```
Correlation_Matrix <- cor(training_data[, -length(training_data)])
corrplot(Correlation_Matrix, type = "lower", tl.cex=0.8)
```

##Compute random forest with PCA to remove correlation

```
Random_Forest <- randomForest(classe~., data = training_data, preprocessing="pca")
print(Random_Forest)
Random_Forest_Validation= predict(Random_Forest, validation)
```
Call:
 randomForest(formula = classe ~ ., data = training_data, preprocessing = "pca") 
               Type of random forest: classification
                     Number of trees: 500
No. of variables tried at each split: 7

        OOB estimate of  error rate: 0.29%
Confusion matrix:
     A    B    C    D    E  class.error
A 5578    2    0    0    0 0.0003584229
B    9 3785    3    0    0 0.0031603898
C    0   10 3409    3    0 0.0037989480
D    0    0   21 3192    3 0.0074626866
E    0    0    1    4 3602 0.0013861935

##Estimate how well the model has been trained
```
print(("Random Forest Cross Validation"))
print(Random_Forest_Validation)

confusionMatrix(validation$classe, Random_Forest_Validation)
accuracy = confusionMatrix(validation$classe, Random_Forest_Validation)$overall['Accuracy']
acc_percentage = (1-accuracy) * 100
print("Percentage :"); print(round(acc_percentage, digits = 2))
```
Confusion Matrix and Statistics

          Reference
Prediction    A    B    C    D    E
         A 1674    0    0    0    0
         B    0 1139    0    0    0
         C    0    0 1026    0    0
         D    0    0    0  964    0
         E    0    0    0    0 1082

Overall Statistics
                                     
               Accuracy : 1          
                 95% CI : (0.9994, 1)
    No Information Rate : 0.2845     
    P-Value [Acc > NIR] : < 2.2e-16  
                                     
                  Kappa : 1          
 Mcnemar's Test P-Value : NA         

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            1.0000   1.0000   1.0000   1.0000   1.0000
Specificity            1.0000   1.0000   1.0000   1.0000   1.0000
Pos Pred Value         1.0000   1.0000   1.0000   1.0000   1.0000
Neg Pred Value         1.0000   1.0000   1.0000   1.0000   1.0000
Prevalence             0.2845   0.1935   0.1743   0.1638   0.1839
Detection Rate         0.2845   0.1935   0.1743   0.1638   0.1839
Detection Prevalence   0.2845   0.1935   0.1743   0.1638   0.1839
Balanced Accuracy      1.0000   1.0000   1.0000   1.0000   1.0000

The accuracy is very high and this is within the 95% confidence interval

We have sufficient confidence to predict for the 20 data set for the quiz

##Now we test for the 20 data set

```
test_data_set = read.csv("pml-testing.csv", na.strings=c("", "NA"))
test_data_set = test_data_set[8:length(test_data_set)]
test_data_set = test_data_set[,cln == 0]

New_Data_Model = predict(Random_Forest, test_data_set)
print(New_Data_Model)
```

1  2  3  4  5  6  7  8  9 10 11 12 13 14 15 16 17 18 19 20 
 B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 
Levels: A B C D E

##Credits 
The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.


